New cardinality estimation algorithms for HyperLogLog sketches

### 2016-05-28 paper/paper.tex
changed definition of tau and sigma

Inserting all these estimators into \eqref{equ:raw_estimate_with_multiplicities} gives
\begin{equation}
\label{equ:correctedestimator}
\symPoissonRateEstimate 
= 
\frac{\symNumReg^2}{2\log(2)\sum_{\symRegVal=-\infty}^\infty \hat{\symCount}'_\symRegVal 2^{-\symRegVal}}
=
\frac{\symNumReg^2}
{2\log(2)
\left(
\symSmallCorrectionFunc(\symCount_0,\symNumReg) + \sum_{\symRegVal=1}^\symRegRange \symCount_\symRegVal 2^{-\symRegVal} + \symLargeCorrectionFunc(\symCount_{\symRegRange+1},\symNumReg) 2^{-(\symRegRange+1)}
\right)}
\end{equation}
where $\symSmallCorrectionFunc(\symCount_0)$ and $\symLargeCorrectionFunc(\symCount_{\symRegRange+1})$ replace $\symCount_0$ and $\symCount_{\symRegRange+1}$ in the \emph{raw} estimate \eqref{equ:intermediate_estimate}, respectively. The functions $\symSmallCorrectionFunc$ and $\symLargeCorrectionFunc$ are given by
\begin{equation}
\symSmallCorrectionFunc(\symCount,\symNumReg) := \symNumReg
\cdot
\left(
\frac{\symCount}{\symNumReg}
+
\sum_{\symRegVal=1}^\infty
\left(\frac{\symCount}{\symNumReg}\right)^{2^\symRegVal} 2^{\symRegVal-1}
\right)
\end{equation}
and
\begin{equation}
\symLargeCorrectionFunc(\symCount,\symNumReg)
:=
\symNumReg
\cdot
\sum_{\symRegVal=1}^\infty
\left(1-\frac{\symCount_{\symRegRange+1}}{\symNumReg}\right)^{2^{-\symRegVal}}
\cdot
\left(1-\left(1-\frac{\symCount_{\symRegRange+1}}{\symNumReg}\right)^{2^{-\symRegVal}}\right)
\cdot
2^{-(\symRegVal-1)}
\end{equation}

For the special case of linear probabilistic counting with $\symRegRange=0$, we get for the  estimator \eqref{equ:correctedestimator}
\begin{equation}
\symPoissonRateEstimate 
= 
\frac{\symNumReg}{\log(2)\sum_{\symRegVal=-\infty}^\infty
\left(\frac{\symCount_0}{\symNumReg}\right)^{2^\symRegVal}2^\symRegVal}
=
\frac{\symNumReg\log\left(\frac{\symNumReg}{\symCount_0}\right)}{\symPowerSeriesFunc\left(\log_2\left(\log\left(\frac{\symNumReg}{\symCount_0}\right)\right)\right)}
\approx
\symNumReg\log\left(\frac{\symNumReg}{\symCount_0}\right).
\end{equation}
Here we used function $\symPowerSeriesFunc$ defined in \eqref{equ:power_series_function} which is known to be approximately equal to 1. Hence, for this special case we get as expected almost the same estimator as given in \eqref{equ:small_range_estimate}.
\subsection{Estimation algorithm}
see \algoref{alg:corrected_raw_estimation}

>>>>

Inserting all these estimators into \eqref{equ:raw_estimate_with_multiplicities} gives
\begin{equation}
\label{equ:correctedestimator}
\symPoissonRateEstimate 
= 
\frac{\symNumReg^2}{2\log(2)\sum_{\symRegVal=-\infty}^\infty \hat{\symCount}'_\symRegVal 2^{-\symRegVal}}
=
\frac{\symNumReg^2}
{2\log(2)
\left(
\symNumReg\,\symSmallCorrectionFunc(\symCount_0/\symNumReg) + \sum_{\symRegVal=1}^\symRegRange \symCount_\symRegVal 2^{-\symRegVal} + \symNumReg\, \symLargeCorrectionFunc(1-\symCount_{\symRegRange+1}/\symNumReg) 2^{-(\symRegRange+1)}
\right)}
\end{equation}
where $\symNumReg\,\symSmallCorrectionFunc(\symCount_0/\symNumReg)$ and $\symNumReg\,\symLargeCorrectionFunc(1-\symCount_{\symRegRange+1}/\symNumReg)$ replace $\symCount_0$ and $\symCount_{\symRegRange+1}$ in the \emph{raw} estimate \eqref{equ:intermediate_estimate}, respectively. The functions $\symSmallCorrectionFunc$ and $\symLargeCorrectionFunc$ are defined as
\begin{equation}
\symSmallCorrectionFunc(\symX) := 
\symX
+
\sum_{\symRegVal=1}^\infty
\symX^{2^\symRegVal} 2^{\symRegVal-1}
\end{equation}
and
\begin{equation}
\symLargeCorrectionFunc(\symX)
:=
\sum_{\symRegVal=1}^\infty
\symX^{2^{-\symRegVal}}
\cdot
\left(1-\symX^{2^{-\symRegVal}}\right)
\cdot
2^{-(\symRegVal-1)}.
\end{equation}

Note that, $2\symSmallCorrectionFunc(\symX) + \symLargeCorrectionFunc(\symX) = -\symPowerSeriesFunc(\log_2(-\log(\symX)))/\left(\log(2)\log(\symX)\right)$. Therefore, we get for the special case of linear probabilistic counting with $\symRegRange=0$
\begin{equation}
\symPoissonRateEstimate 
= 
\frac{\symNumReg}{\log(2)\left(
2\symSmallCorrectionFunc(\symCount_0/\symNumReg)
+
\symLargeCorrectionFunc(\symCount_0/\symNumReg)
\right)}
=
\frac{\symNumReg\log\left(\frac{\symNumReg}{\symCount_0}\right)}{\symPowerSeriesFunc\left(\log_2\left(\log\left(\frac{\symNumReg}{\symCount_0}\right)\right)\right)}
\approx
\symNumReg\log\left(\frac{\symNumReg}{\symCount_0}\right).
\end{equation}
almost the same estimator as given in \eqref{equ:small_range_estimate}. Here we used that $\symPowerSeriesFunc$ is known to be approximately equal to 1.
\subsection{Estimation algorithm}
see \algoref{alg:corrected_raw_estimation}
